LFQ: Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning

### 2020-10-27 paper/presentation_lcn.tex
Final version of the LCN presentation

\begin{frame}{``Bufferbloat''}
  \begin{itemize}
  \item If queues in routers/switches too small: \textbf{Underutilization}
  \item \textbf{Solution:} Make queues very \textbf{large} for maximum throughput!
  \item \textbf{New problem:} Packets \textbf{wait} a long time (several seconds) in the queue: Bufferbloat
  \end{itemize}
\end{frame}

\begin{frame}{CoDel}{Active Queue Management against Bufferbloat}
  \begin{itemize}
  \item Moving time window of 100\,ms
  \item Queuing Delay \textbf{must be $<$ 5\,ms} once in each window
  \item Otherwise: Drop packet(s)
  \end{itemize}
\end{frame}

@@

\begin{frame}{\texttt{fq\_codel}}{Combining fair queuing with CoDel}
  \begin{itemize}
  \item State-of-the-art
  \item Separate queues
  \item Each queue managed by CoDel
  \item Keeps each flow's queue $<$ 5\,ms
  \item Linux implementation: \texttt{fq\_codel}
  \pause
  \end{itemize}
  \begin{alertblock}{Inadequate interaction with Congestion Control:}
	\textit{Cubic} doesn't achieve full throughput or keeps standing queue 
  \end{alertblock}
\end{frame}

@@

\begin{frame}{\texttt{cocoa} qdisc\footnote{\textit{Cocoa: Congestion Control Aware Queuing}, Bachl, Fabini, Zseby, 2020}}
\begin{itemize}
\item Fair queuing-based qdisc which \textbf{adapts} buffer \textbf{depending on congestion control}
\item Manages to achieve optimal throughput while keeping \textbf{delay} from buffering \textbf{minimal}
\item \textbf{Works} for common congestion controls
\end{itemize}
\pause
\begin{alertblock}{Hand-crafted algorithm:}
Works well for current congestion controls but might fail badly for new congestion controls.
\end{alertblock}
\end{frame}

@@

\begin{frame}{Deep Learning for Buffering}
\begin{itemize}
\item Have a separate queue for each flow in a switch/router (fair queuing)
\item Define some reward function (like high throughput, low delay)
\item Use \textbf{Deep Learning}
\begin{itemize} \item to \textbf{fingerprint} each flow
\item and to use the \textbf{correct buffer size} for the flow \textbf{based on experience}, which \textbf{maximizes the reward}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Reward}
\begin{align*}
\textit{Reward} = \textit{bandwidth}-\alpha\times\textit{queue size}
\end{align*}
\end{frame}

@@

\begin{frame}{Input features 2}
\begin{itemize}
\item Don't use these features themselves but exponentially weighted moving averages
\item Advantage: No need to keep data around except for features themselves
\end{itemize}
\end{frame}

@@

\begin{frame}{Implementation}
\begin{itemize}
\item ns-3 for network simulation
\item Pytorch's C++ API for Deep Learning
\item Integrate Pytorch into ns-3
\item Develop queuing discipline based on fair queuing
\item Each flow is managed by Reinforcement Learning based on Deep Learning
\end{itemize}
\end{frame}

\section{Offline Learning}
\begin{frame}{Offline Learning}
\begin{itemize}
\item Randomly sample bandwidth, delay, congestion control and flow duration
\item For each flow, at random time, launch experiment
\item Experiment is A/B testing: 
\begin{itemize}
\item Continue with current buffer size +1 and current buffer size -1 (two different experiment)
\item Let the neural network learn buffer size that performed better regarding reward
\end{itemize}
\item Train using L1 loss (mean absolute error), using a couple of results together as a batch
\end{itemize}
Training takes several hundred thousand flows to converge. 
\end{frame}

@@

\begin{frame}{Visualizing success}
Should learn
\begin{itemize}
\item Larger bandwidth $\rightarrow$ larger buffer
\item Larger delay $\rightarrow$ larger buffer
\end{itemize}
\end{frame}

@@

\begin{frame}{New Reno, varying bandwidth}
\centering
\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_bandwidth"}.pdf}
\end{frame}

\begin{frame}{New Reno, varying delay}
\centering
\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_delay"}.pdf}
\end{frame}

\begin{frame}{Bic, varying bandwidth}
\centering
\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
\end{frame}

\begin{frame}{Bic, varying delay}
\centering
\includegraphics[width=0.9\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
\end{frame}

@@

\begin{frame}{Correlations}
\begin{table}[h]
\caption{Correlation between bandwidth/delay of the link and output buffer size.} \label{tab:corrOfflineSmallAlpha}
\centering
\begin{tabular}{lrr} \toprule
Congestion control & bandwidth & delay \\ \midrule
New Reno & 99.3\% & 96.5\% \\
BIC & 75.4\% & 62\% \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

@@

\begin{frame}{Conclusion}
\begin{itemize}
\item LFQ is based on fair queuing
\item It fingerprints each flow
\item It learns to optimize a reward function
\item It achieves high throughput and low delay when compared to competing solutions
\item It has low computational overhead
\item We envision deployment close to end users (on switches, routers)
\end{itemize}
\end{frame}

>>>>

@@

%\begin{frame}{``Bufferbloat''}
%  \begin{itemize}
%  \item If queues in routers/switches too small: \textbf{Underutilization}
%  \item \textbf{Solution:} Make queues very \textbf{large} for maximum throughput!
%  \item \textbf{New problem:} Packets \textbf{wait} a long time (several seconds) in the queue: Bufferbloat
%  \end{itemize}
%\end{frame}

%\begin{frame}{CoDel}{Active Queue Management against Bufferbloat}
%  \begin{itemize}
%  \item Moving time window of 100\,ms
%  \item Queuing Delay \textbf{must be $<$ 5\,ms} once in each window
%  \item Otherwise: Drop packet(s)
%  \end{itemize}
%\end{frame}

%\begin{frame}{\texttt{fq\_codel}}{Combining fair queuing with CoDel}
%  \begin{itemize}
%%  \item State-of-the-art
%  \item Separate queues
%  \item Each queue managed by CoDel
%  \item Keeps each flow's queue $<$ 5\,ms
%  \item Linux implementation: \texttt{fq\_codel}
%  \pause
%  \end{itemize}
%  \begin{alertblock}{Inadequate interaction with Congestion Control:}
%	\textit{Cubic} doesn't achieve full throughput or keeps standing queue 
%  \end{alertblock}
%\end{frame}

@@

%\begin{frame}{\texttt{cocoa} qdisc\footnote{\textit{Cocoa: Congestion Control Aware Queuing}, Bachl, Fabini, Zseby, 2020}}
%\begin{itemize}
%\item Fair queuing-based qdisc which \textbf{adapts} buffer \textbf{depending on congestion control}
%\item Manages to achieve optimal throughput while keeping \textbf{delay} from buffering \textbf{minimal}
%\item \textbf{Works} for common congestion controls
%\end{itemize}
%\pause
%\begin{alertblock}{Hand-crafted algorithm:}
%Works well for current congestion controls but might fail badly for new congestion controls.
%\end{alertblock}
%\end{frame}

@@

\begin{frame}{Deep Reinforcement Learning for Buffering}
\begin{itemize}
\item Have a separate queue for each flow in a switch/router (fair queuing)
\item Define some reward function (like high throughput, low delay)
\item Use \textbf{Deep Reinforcement Learning}
\begin{itemize} \item to \textbf{fingerprint} each flow
\item and to use the \textbf{correct buffer size} for the flow \textbf{based on experience}, which \textbf{maximizes the reward}
\end{itemize}
\item We call it \textit{Learning Fair Queuing}
\end{itemize}
\end{frame}

\begin{frame}{Reward}
\begin{align*}
\textit{Reward} = \textit{throughput}-\alpha\times\textit{queue size}
\end{align*}
\end{frame}

@@

%\begin{frame}{Input features 2}
%\begin{itemize}
%\item Don't use these features themselves but exponentially weighted moving averages
%\item Advantage: No need to keep data around except for features themselves
%\end{itemize}
%\end{frame}

@@

\begin{frame}{Implementation}
\begin{itemize}
\item ns-3 for network simulation
\item Pytorch's C++ API for Deep Learning
\item Integrate Pytorch into ns-3
\item Develop queuing discipline based on fair queuing
\item Each flow is managed by Deep Reinforcement Learning
\end{itemize}
\end{frame}

\section{Offline Learning}
\begin{frame}{Offline Learning}
\begin{itemize}
\item Randomly sample bandwidth, delay, congestion control and flow duration
\item Simulate each flow continuously using the optimal buffer size output by the Deep Reinforcement Learning
\item For each flow, at a random time, launch experiment (A/B testing):
\begin{itemize}
\item Continue one experiment with current buffer size +1 packet and one with current buffer size -1 packet (two different experiments)
\end{itemize}
\item Let the neural network learn which buffer size achieved the higher reward
\item Train using L1 loss (mean absolute error), using a couple of results together as a batch
\end{itemize}
Training takes several hundred thousand flows to converge. 
\end{frame}

@@

\begin{frame}{Measuring success}
Should learn
\begin{itemize}
\item Larger bandwidth $\rightarrow$ larger buffer
\item Larger delay $\rightarrow$ larger buffer
\item New Reno Congestion Control should get bigger buffer than BIC
\end{itemize}
\end{frame}

@@

%\begin{frame}{New Reno, varying bandwidth}
%\centering
%\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_bandwidth"}.pdf}
%\end{frame}
%
%\begin{frame}{New Reno, varying delay}
%\centering
%\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_delay"}.pdf}
%\end{frame}
%
%\begin{frame}{Bic, varying bandwidth}
%\centering
%\includegraphics[width=0.85\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
%\end{frame}
%
%\begin{frame}{Bic, varying delay}
%\centering
%\includegraphics[width=0.9\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
%\end{frame}

@@

%\begin{frame}{Correlations}
%\begin{table}[h]
%\caption{Correlation between bandwidth/delay of the link and output buffer size.} \label{tab:corrOfflineSmallAlpha}
%\centering
%\begin{tabular}{lrr} \toprule
%Congestion control & bandwidth & delay \\ \midrule
%New Reno & 99.3\% & 96.5\% \\
%BIC & 75.4\% & 62\% \\
%\bottomrule
%\end{tabular}
%\end{table}
%\end{frame}

@@

\begin{frame}{Conclusion}
\begin{itemize}
\item LFQ is based on fair queuing
\item It fingerprints each flow
\item It learns to optimize a reward function
\item It achieves high throughput and low delay when compared to competing solutions
\item It has low computational overhead
\item We envision deployment close to end users (for example on home routers)
\end{itemize}
\end{frame}
